Speakers of diverse languages structure their utterances for efficient communication

### 2019-11-06 Paper/info-curves.Rmd
methods edits on paper

@shannon1948 defined information as "the reduction in uncertainty about one variable given that the value of another variable is known". We use a metric proposed by Shannon and applied by @levy2008: lexical surprisal, which defines the information in word based on the ratio of possible continuations of the sentence after to before the word is seen. Equivalently, we can compute surprisal with the predictability of the word, as in the formula below. The surprisal of a word is inversely proportional to the predictability of a word, such that less common and less predictable words carry more information.

$$surprisal(word) = -log P(word)$$

@@

However, when reading or listening, people don’t just consider each word as an isolated linguistic signal. Instead, listeners use the words they have already heard to predict and decode the word they are currently processing. Following this incremental processing paradigm, we can also condition the surprisal of a word in its context. In the formula below, widenotes the word currently being read or heard, while wi-1denotes the first word before the current word, wi-2denotes the second word before the current word, and so on.

$$surprisal(w_i|w_{i-1}w_{i-2}...) = -log P(w_i|w_{i-1}w_{i-2}...)$$
$$= -log \frac{P(w_i,w_{i-1}w_{i-2},...)}{P(w_{i-1}w_{i-2}...)}$$

@@

We train an ngram language model on a given corpus with order set to $1$ for the frequency-model, $2$ for a bigram model and $3$ for a trigram model. An ngram language model produces a probability for each continuation of each $n-1$ words it receives as input. Models with order greater than $3$ have issues with overfitting to the corpus, while bigram models generally underperform trigram models. In our contextual models, we face another issue of overfitting: we only train our model on those utterances which occur in the corpus and test our model on the same utterances. This ignores possible other utterances which the speakers could have produced, e.g. the words "I", "saw" and "bears" are in the corpus vocabulary, which the speaker may not have produced in the corpus but could have produced. To combat this issue, we use modified Kneser-Ney smoothing as implemented in the KenLM toolkit [@heafield2013scalable]. Briefly, this smoothing technique discounts all ngram counts, which reduces the impact of rare ngrams on probability calculations, and interpolates lower-order ngrams into the calcuations. These lower-order ngrams are weighted according to the number of distinct contexts they occur as a continuation (e.g. "Francisco" may be a common word in a corpus, but likely only occurs after "San" as in "San Francisco", so it receives a lower weighting). For a full explanation of modified Kneser-Ney smoothing, see @chen1999empirical.

@@

We first turn to working with written English in the British National Corpus [BNC; @leech1992]. The BNC is a collection of spoken and written records (90% written) from the turn of the century, intended to be a representative sample of British English. Using their word entropy metric without context, Yu et al. found a distinctive three-step distribution for information in written English sentences in the corpus. The first word tended to contain little information. While the middle words of sentences each had more information than the first word, information was flat and non-increasing across the middle of sentences. The final word, they found, contained the most information out of any in the sentence, with a noticeable spike in information. They found the same distribution across sentence lengths, from sentences with 15 words to sentences with 45 words.

>>>>

@shannon1948 defined information as "the reduction in uncertainty about one variable given that the value of another variable is known". We use a metric proposed for the study of information transmission more generally by Shannon and applied to words specifically by @levy2008: lexical surprisal. This measure defines the information in word based on the ratio of possible continuations of the sentence after to before the word is seen. Equivalently, we can compute surprisal with the predictability of the word, as in the formula below. The surprisal of a word is inversely proportional to the predictability of a word, such that less common and less predictable words carry more information.

$$\text{surprisal}(\text{word}) = -\log P(\text{word})$$

@@

However, when reading or listening, people don’t just consider each word as an isolated linguistic signal. Instead, listeners use the words they have already heard to predict and decode the current word. Following this incremental processing paradigm, we can also condition the surprisal of a word in its context. In the formula below, widenotes the word currently being read or heard, while $w_{i-1}$ denotes the first word before the current word, $w_{i-2}$ denotes the second word before the current word, and so on.

$$\text{surprisal}(w_i|w_{i-1}w_{i-2}...) = -log P(w_i|w_{i-1}w_{i-2}...)$$
$$= -log \frac{P(w_i,w_{i-1}w_{i-2},...)}{P(w_{i-1}w_{i-2}...)}$$

@@

Ideally, we would like to predict each word using all of the information available. For example, in an utterance of twenty words, we would be able to use the previous $19$ words of context to predict the twentieth word. However, we would need to train on a corpus of many trillion word tokens to predict with this amount of context, and we want to directly compare how predictable each word is regardless of its position in an utterance. Therefore, we use the *Markov assumption*, conditioning our language models' predictions on only a few words of preceding context at most. We train ngram language model on each corpus.

$$\text{surprisal}(w_i|w_{i-1}w_{i-2}...) \approx \text{surprisal}(w_i|w_{i-1}w_{i-2})$$

For our frequency-based model, we don't incorporate context and so each ngram has an order of $1$: unigrams. To incorporate context into our models, we train bigram and trigram language models, which incorporate one and two words of context for each processed word, respectively. Although these models may seem to use an inconsequential amount of context when predicting the next word, bigram and trigram models introduce a great deal of improvement over unigram models across tasks [@chen1999empirical]. Models with order greater than $3$ have issues with overfitting to the corpus and only predicting observed sequences, often generalizing poorly.

In our contextual models, we face another issue of overfitting: we only train our model on those utterances which occur in the corpus and test our model on the same utterances. This ignores possible other utterances which the speakers could have produced, e.g. the words "I", "saw" and "bears" are in the corpus vocabulary, which the speaker may not have produced in the corpus but could have produced. To combat this issue, we use modified Kneser-Ney smoothing as implemented in the KenLM toolkit [@heafield2013scalable]. Briefly, this smoothing technique discounts all ngram counts, which reduces the impact of rare ngrams on probability calculations, and interpolates lower-order ngrams into the calcuations. These lower-order ngrams are weighted according to the number of distinct contexts they occur as a continuation (e.g. "Francisco" may be a common word in a corpus, but likely only occurs after "San" as in "San Francisco", so it receives a lower weighting). For a more complete explanation of modified Kneser-Ney smoothing, see @chen1999empirical.

@@

We first turn to working with written English in the British National Corpus [BNC; @leech1992]. The BNC is a collection of spoken and written records (90% written) from the turn of the century, intended to be a representative sample of British English. Using their word entropy metric without context, Yu et al. found a distinctive three-step distribution for information in written English sentences in the corpus. The first word tended to contain little information. While the middle words of sentences each had more information than the first word, they found a flat and non-increasing rate of information transmission across the middle of sentences. The final word contained the most, though not most, of the information out of any in the sentence, with a noticeable spike in information. They found the same distribution across sentence lengths, from sentences with $15$ words to sentences with $45$ words.
